---
title: "Tidy Data"
output: github_document
---

```{r}
library(tidyverse)
```

## PULSE data with `pivot_longer`

```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, ##these are the columns I want to change
    names_to = "obs_time", ##this is what I want the new column containing the new variable to be called
    values_to = "bdi_score", ##this is what I want the existing values to be stored as
    names_prefix = "bdi_score_" ##this will help me get rid of unecessary prefixes in the names column
  ) |> 
  mutate(
    obs_time = replace(obs_time, obs_time == "bl", "00m")
  ) ##I want to replace the "bl" value with "00m" for consistency

```


## Learning assessment
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(litter_number, gd0_weight, gd18_weight) |> 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight"
  ) |> 
  mutate(
    gd = case_match(
      gd,
      "gd0_weight" ~ 0,
      "gd18_weight" ~18
    ) ##case_match is good for replacing multiple values (not just one-offs)
  )
```

To answer, is this version tidy? - it depends! What are you trying to do with the data after? If we wanted to create a weight gain variable, it would be harder to do it this way. Form should follow function!

## LOTR data

This is going to be a bit trickier because there are three separate tables in the excel sheet.

```{r}
lotr_fellowship_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship")

lotr_two_towers_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "F3:H6") |> 
  mutate(movie = "two towers")

lotr_return_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "J3:L6") |> 
  mutate(movie = "return of the king")

lotr_df = 
  bind_rows(lotr_fellowship_df, lotr_two_towers_df, lotr_return_df) |> 
  pivot_longer(
    Female:Male,
    names_to = "gender",
    values_to = "word"
    ) |> 
  relocate(
    movie
    )
```

## Revisit FAS to join datasets

```{r}
litters_df_2 = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  mutate(wt_gain = gd18_weight - gd0_weight) |> 
  select(litter_number, group, wt_gain) |> 
  separate(group, into = c("dose", "day_of_tx"), 3) ##this lets me separate the column into two separate variables. 

pups_df_2 = 
  read_csv("data/FAS_pups.csv") |> 
  janitor::clean_names() |> 
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  ) ##always code your categorical using the actual levels, not numeric- because you shouldn't ever assume your downstream user knows what the numeric corresponds to

fas_df = 
  left_join(pups_df_2, litters_df_2, by = "litter_number")
```

