---
title: "Simple document"
output: github_document
---

## Loading the data!

```{r}
library(tidyverse)
library(readxl)
library(haven)
```


Let's import the `FAS_litters.csv` using a relative path. 

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

litters_df = 
  janitor::clean_names(litters_df)
#inside the janitor package, I can just load the function I need to have (and not all the others)
```

Let's import the `FAS_litters.csv` using an absolute path. 

```{r}
litters_df_absolute = 
  read.csv("~/Desktop/P8105/data_wrangling_i/data/FAS_litters.csv")

litters_df_absolute = 
  janitor::clean_names(litters_df_absolute)
```

Now we're going to load in the `FAS_pups.csv` dataset. 

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df = 
  janitor::clean_names(pups_df)
```

## Looking at the data!!

```{r}
litters_df

head(litters_df)

tail(litters_df)
```

To see the whole dataset, use `View`. 

```{r, eval = FALSE}
view(litters_df) #don't include in your R markdown files!!! very big. 
```

Let's look at some data summaries.
```{r}
str(litters_df)

skimr::skim(litters_df)
```

##Options in read_*

What if we need to skip over some rows of data? 

```{r}
litters_df_2 = 
  read_csv(
    "data/FAS_litters.csv",
    skip = 10, col_names = FALSE)
```


Let's look at some NA values- if we know a specific value is used to represent missing data, we can specify that when we load the data.
```{r}
litters_df_3 = 
  read_csv("data/FAS_litters.csv",
           na = c("NA", 19))
```

Specifying column types

```{r}
litters_df_4 = 
  read_csv(
    "data/FAS_litters.csv", 
    col_types = cols(
      Group = col_factor(),
      `GD0 weight` = col_character()
    ))
```

##Other file types

Import and XLSX file first. Readxl is SUPER similar to read_csv and developed by the same people, so all the same principles apply!

```{r}
mlb_df = 
  read_excel("data/mlb11.xlsx")
```

Now we're going to import a SAS file. You need the haven library. 

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")
```

