---
title: "Data manipulation"
output: github_document
---

## Loading the data!

```{r}
library(tidyverse)
```

Load FAS litters and pups

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## `select` function

`Select` is used to select columns.

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, group, gd0_weight:gd_of_birth) ##everything up to and including these columns

select(litters_df, group, starts_with("pups")) ##selects all that begin with pups! also have ends with and contains.

select(litters_df, -litter_number) ## - means "not this"

select(litters_df, group, litter_id = litter_number) ##select and replace the name of a column

select(litters_df, group, litter_id = litter_number, everything()) ## lets you keep all the columns, good for reordering the columns 

rename(litters_df, litter_id = litter_number) ##amother way to rename a column and keep all the others

relocate(litters_df, litter_number)

```

What's the difference between `select` and `pull`?

```{r}
select(litters_df, group) ##preserves the data in a dataframe!!!
pull(litters_df, group) ##data no longer in a dataframe, but still adheres to tidyverse logic
```

## Learning assessment 

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` function

Get rid of rows using `filter`

```{r}
filter(litters_df, group == "Con7")

filter(litters_df, group != "Con8")

filter(litters_df, gd0_weight > 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8")) ##this and the previous accomplish the same thing

filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight >20) ##this sequentially adds an additional filter to the results pulled from the first part 
```

What about getting rid of NAs? Don't use `filter`, use `drop_na`

```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```


## `mutate` function

This is used to change variables.

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, group = str_to_lower(group))

mutate(litters_df,
       wt_gain = gd18_weight - gd0_weight, 
       group = str_to_lower(group)) ##can string these together with commas
```

## `arrange` function

```{r}
arrange(litters_df, gd0_weight)
arrange(litters_df, group, gd0_weight) ##order that you want the varaibles prioritized for arranging!
```

## pipes


Shortcut for pipes is command shift M


```{r}
litters_df_2 = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(-starts_with("pups")) |> 
  mutate(
    group = str_to_lower(group), 
    wt_gain = gd18_weight - gd0_weight,
  ) |> 
  drop_na(wt_gain) |> 
  arrange(group, wt_gain) ##this makes all the changes we did individually above!!


pups_df_2 = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names()


```

